---
title: "Vaccine Decision Making: 2016 v 2018"
author: "K. Anderson"
date: "2024-05-15"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    code_folding: hide
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

**Package and data loading**

```{r}
library(haven)
library(tidyverse)
library(car)
library(Hmisc)
library(GGally)
library(ltm)
library(survey)

# Import Data -----------------------------------------------------------------------------------

# Temporary Data folder in this repo##
# Data are available upon request.
# SurveyData2016 are located at 
# https://github.com/PandemicSystems/DP3-survey/blob/main/data/2016%20Data%20Files%20(11-9-2016)/UnivGeorgia_AdultVacc2016Main.sav
# SurveyData2018 are located at 
# https://github.com/PandemicSystems/DP3-survey/blob/main/data/2018%20Vacc%20Data%20Files%20(11-16-18)/pn8351_ugavacc_111618_unweighted.sas7bdat
# SurveyData2016 <- read_sav("temp_data/UnivGeorgia_AdultVacc2016Main.sav")
# SurveyData2018 <- read_sas("temp_data/pn8351_ugavacc_111618_unweighted.sas7bdat")

```

# 2016 Survey Analysis

**Independent Variables**\
*Vaccine Confidence*:\
Q20_A : <u>All</u> of the vaccines recommended for adults are safe.\
Q20_B : <u>All</u> of the vaccines recommended for adults are effective.<br>
Q20_C : Your overall health will benefit from getting <u>all</u> the recommended vaccines for adults.\
Composite Score: Q20_Avg

*Vaccine Hesitancy*\
Q19: Overall, how hesitant about getting recommended adult vaccines would you consider yourself to be?\
*Vaccine History*\
Q8B : Thinking about 2013, 2014, and 2015, in how many of those years did you get a flu vaccination?\
*Shot Likelihood*\
Q9: How likely are you to get a flu vaccination in this flu season – that is, this Fall or Winter?\

**Block 1: Demographics**\
GENDER + AGE + RACETHNICITY + INCOME + EDUC\

**Block 2: Information**\
*Perceived Flu Knowledge*\
 Q32 :How well informed would you say you are about each of the following:\
 A.	Seasonal influenza or flu?\
 B.	Who should receive the seasonal influenza or flu vaccine?\
 C.	The benefits associated with getting seasonal influenza or flu vaccine?\
 D.	The risks associated with not getting a seasonal influenza or flu vaccine?\
 Composite Score: Q32_Avg
*Dr. Recommendation*\
Q7: In the past year, has a physician or healthcare provider told you that you should receive an influenza or flu vaccine?\

**Block 3: Flu Impacts**\
*Flu Seriousness*\
Q37: In general, if you get the flu, how serious do you think the illness would be <u>for you personally</u>?\
*Flu Spread*\
Q39: In general, if you get the flu, how likely do you think it is that you could pass the flu on to someone else?\

**Block 4: Vaccine Attitudes (with Vaccine Confidence)**\
*Vaccine Hesitancy*\
Q19: Overall, how hesitant about getting recommended adult vaccines would you consider yourself to be?\
*Vaccine Difficulty*\
Q35: In general, how easy or difficult is it – or would it be – for you to get a flu shot or vaccination if you wanted one?\


**Block 5: Health Decision-Making Preferences**\
*Health Decision-Making Preferences*\
Q16_A \
(a) “I prefer natural ways to prevent and treat illness” vs. “I prefer mainstream medicine approaches to prevent and treat illness,”\
Q16_E *(Q16_F in 2018 Survey)*\
(b) “I prefer complementary or alternative medicine approaches to staying healthy and treating disease” vs. “I prefer science and evidence-based medicine approaches to staying healthy and treating illnesses,”	\
Q16_B\
(c) “I am most interested in the risks and potential side effects of a recommended drug or medicine.” vs. “I am most interested in the benefits and value of a recommended drug or medicine,”	\
Q16_F *(Q16_G in 2018 Survey)*\
(d) “I usually doubt health and medical advice from doctors and experts” vs. “I usually trust health and medical advice from doctors and experts,”\	
Q16_D *(Q16_E in 2018 Survey)*\
(e) “When sick or ill, I prefer as little medicine or treatment as possible” vs. “When sick or ill, I prefer as much medicine or treatment as possible,”	\
Q16_H *(Q16_I in 2018 Survey)*\
(f) “I usually wait until I have an illness or health problem and then take steps or actions to treat it” vs. “I usually take steps or actions in advance to prevent an illness or health problem.”\
Composite Score: Q16_Avg


## 2016 Dataframe including both all components for the idependent and independent variables
```{r}

Survey_ICA_16 <- SurveyData2016 %>% 
dplyr::select(CaseId, WEIGHT, AGE, GENDER, RACETHNICITY, INCOME, EDUC, Q20_A, Q20_B, Q20_C, Q8B, Q9, Q32_A, Q32_B, Q32_C, Q32_D, Q7, Q37, Q39, Q35, Q19, Q16_A, Q16_E, Q16_B, Q16_F, Q16_D, Q16_H )%>%
#"Missing" and "I don't know" indicators (98, 99, 77) converted to "NAs" which were then kept in the data set. This condition is not applied to AGE
  mutate(across(-AGE, ~ na_if(., 98) %>% na_if(99) %>% na_if(77))) %>%
 #Reordering and/or reassigning response values such that higher values ~ higher behavior likelihood
  mutate(RACETHNICITY = case_when(RACETHNICITY > 1 ~ "0", TRUE ~ "1")) %>%
  mutate_at(vars(Q8B), ~.-1)%>%
  mutate(Q9 = case_when(Q9 == 1 ~ 4, Q9 == 2 ~ 3, Q9 == 3 ~ 2, Q9 == 4 ~ 1, Q9 == 5 ~ 5, TRUE ~ NA_real_))%>% 
  mutate(Q37 = case_when(Q37 == 1 ~ 4, Q37 == 2 ~ 3, Q37 == 3 ~ 2, Q37 == 4 ~ 1, TRUE ~ NA_real_))%>%
  mutate(Q39 = case_when(Q39 == 1 ~ 4, Q39 == 2 ~ 3, Q39 == 3 ~ 2, Q39 == 4 ~ 1, TRUE ~ NA_real_))%>%
  mutate_at(vars(RACETHNICITY), as.numeric) %>%
  rowwise() %>%
#Calculating a score (single value) for specific categories 
  mutate(Q32_Avg = mean(c(Q32_A, Q32_B, Q32_C, Q32_D), na.rm = TRUE)) %>% # Flu Knowledge
  mutate(Q20_Avg = mean(c(Q20_A, Q20_B, Q20_C), na.rm = TRUE)) %>% # Vaccine Confidence
  mutate(Q16_Avg = mean(c(Q16_A, Q16_E, Q16_B, Q16_F, Q16_D, Q16_H), na.rm = TRUE)) %>% #Health decision-making preference
  ungroup()

```

## 2016 Regression Model
```{r}

#Regression model with all variables
model_all_16 <- lm(Q9 ~ GENDER + AGE + RACETHNICITY + INCOME + EDUC + Q32_Avg + Q7  + Q37 + Q39 + Q20_Avg + Q19 + Q35 + Q16_Avg, data = na.exclude(Survey_ICA_16),weight = WEIGHT)
summary(model_all_16)

#To Calculate Block 1 R-squared and significance
model_B1_16 <- lm(Q9 ~ GENDER + AGE + RACETHNICITY + INCOME + EDUC, data = na.exclude(Survey_ICA_16), weight = WEIGHT)
summary(model_B1_16)

#Calculating table column sample size: 
setNames(c(sum(complete.cases(Survey_ICA_16$Q20_Avg)) ,sum(complete.cases(Survey_ICA_16$Q19)), sum(complete.cases(Survey_ICA_16$Q8B)), sum(complete.cases(Survey_ICA_16$Q9))), c("Conf.", "Hes.", "Past Vx", "Intent"))

```

## 2016 Incremental $R^2$
This script below is used to calculate the incremental $R^2$ for the hierarchical ordinary least squares (OLS) (adding variables by Blocks) regression models for the prediction of vaccine confidence, vaccine hesitancy, vaccine history and vaccination likelihood. 

```{r}

data_16 = na.exclude(Survey_ICA_16)

extensions_16 <- list(c("Q32_Avg", "Q7"), c("Q37", "Q39"), c("Q20_Avg", "Q19", "Q35"), "Q16_Avg")#Each c() is a Block 2+
response_16 <- list("Q20_Avg", "Q19", "Q8B", "Q9" ) #Response Variables: Confidence, Hesitancy, Past Vax, Future Vax

results_list_16 <- list()

for (i in seq_along(response_16)){

base_formula_16 <- as.formula(paste(response_16[[i]], "~ GENDER + AGE + RACETHNICITY + INCOME + EDUC")) #Block 1
base_model_16 <- lm(base_formula_16, data = data_16, weights = WEIGHT)
base_r_squared_16 <- summary(base_model_16)$r.squared

results_list_16[[i]] <- list()

for (ii in seq_along(extensions_16)) {
    # Combine the base formula and the current extension
    extended_formula_16 <- update(base_formula_16, paste("~ . +", paste(extensions_16[[ii]], collapse = " + ")))
    
    # Fit the extended model
    extended_model_16 <- lm(extended_formula_16, data = data_16, weights = WEIGHT)

    # Get the R-squared for the extended model
    extended_r_squared_16 <- summary(extended_model_16)$r.squared

    # Calculate the incremental R-squared
    delta_r_squared_16 <- extended_r_squared_16 - base_r_squared_16

    #Likelihood Ratio Test
    LRT_16 <- anova(base_model_16, extended_model_16, test = "LRT")
    
    # Result List
    results_list_16[[i]][[ii]] <- list(
      base_r_squared = base_r_squared_16,
      extended_formula = extended_formula_16,
      extended_model_sumry = summary(extended_model_16),
      extended_r_squared = extended_r_squared_16,
      delta_r_squared = delta_r_squared_16,
      LRT = LRT_16)
    
    # Resetting base formula to last extension
    base_formula_16 <- extended_formula_16
    base_model_16 <- lm(base_formula_16, data = data_16, weights = WEIGHT)
    base_r_squared_16 <- summary(base_model_16)$r.squared
  }

} 

print(results_list_16)

```

# 2018 Survey Analysis

## 2018 Dataframe including both all components for the idependent and independent variables

```{r}

 Survey_ICA_18 <- SurveyData2018 %>% 
  dplyr::select(CaseId,GENDER, AGE, RACETHNICITY, INCOME, EDUC, Q20_A, Q20_B, Q20_C, Q8B, Q9, Q32_A, Q32_B, Q32_C, Q32_D, Q7, Q37, Q39, Q35, Q19, Q16_A, Q16_F, Q16_B, Q16_G, Q16_E, Q16_I )%>%
#"Missing" and "I don't know" indicators (98, 99, 77) converted to "NAs" which were then kept in the data set.
  mutate(across(-AGE, ~ na_if(., 98) %>% na_if(99) %>% na_if(77))) %>%
#Reordering and/or reassigning response values such that higher values ~ higher behavior likelihood
  mutate(RACETHNICITY = case_when(RACETHNICITY > 1 ~ "0", TRUE ~ "1")) %>%
  mutate_at(vars(Q8B), ~.-1)%>%
  mutate(Q9 = case_when(Q9 == 1 ~ 4, Q9 == 2 ~ 3, Q9 == 3 ~ 2, Q9 == 4 ~ 1, Q9 == 5 ~ 5, TRUE ~ NA_real_))%>% #"Other"
  mutate(Q37 = case_when(Q37 == 1 ~ 4, Q37 == 2 ~ 3, Q37 == 3 ~ 2, Q37 == 4 ~ 1, TRUE ~ NA_real_))%>%
  mutate(Q39 = case_when(Q39 == 1 ~ 4, Q39 == 2 ~ 3, Q39 == 3 ~ 2, Q39 == 4 ~ 1, TRUE ~ NA_real_))%>%
#Ensuring values are numeric
  mutate_at(vars(RACETHNICITY), as.numeric)%>%
  rowwise() %>%
#Calculating a score (single value) for specific categories 
  mutate(Q32_Avg = mean(c(Q32_A, Q32_B, Q32_C, Q32_D), na.rm = TRUE))%>% # Flu Knowledge
  mutate(Q20_Avg = mean(c(Q20_A, Q20_B, Q20_C), na.rm = TRUE)) %>% # Vaccine Confidence
  mutate(Q16_Avg = mean(c(Q16_A, Q16_F, Q16_B, Q16_G, Q16_E, Q16_I), na.rm = TRUE)) %>% #Health decision-making preference
  ungroup()
```

## 2018 Regression Model
```{r}
#Regression model with all variables
model_all_18 <- lm(Q8B ~ GENDER + AGE + RACETHNICITY + INCOME + EDUC + Q32_Avg + Q7  + Q37 + Q39 + Q20_Avg + Q19 + Q35 + Q16_Avg, data = Survey_ICA_18)
summary(model_all_18)

#For the Block 1 Incremental R-squared significance
model_B1_18 <- lm(Q20_Avg ~ GENDER + AGE + RACETHNICITY + INCOME + EDUC, data = na.exclude(Survey_ICA_18))
summary(model_B1_18)

#Calculating table column sample size
c(sum(complete.cases(Survey_ICA_18$Q20_Avg)), sum(complete.cases(Survey_ICA_18$Q19)), sum(complete.cases(Survey_ICA_18$Q8B)), sum(complete.cases(Survey_ICA_18$Q9)))

```

## 2018 Incremental $R^2$
This script below is used to calculate the incremental $R^2$ for the hierarchical ordinary least squares (OLS) (adding variables by Blocks) regression models for the prediction of vaccine confidence, vaccine hesitancy, vaccine history and vaccination likelihood.

```{r}

data_18 <- na.exclude(Survey_ICA_18)

extensions_18 <- list(c("Q32_Avg", "Q7"), c("Q37", "Q39"), c("Q20_Avg", "Q19", "Q35"), "Q16_Avg")
response_18 <- list("Q20_Avg", "Q19", "Q8B", "Q9" )

results_list_18 <- list()

for (i in seq_along(response_18)){

base_formula_18 <- as.formula(paste(response_18[[i]], "~ GENDER + AGE + RACETHNICITY + INCOME + EDUC"))#Block 1
base_model_18 <- lm(base_formula_18, data = data_18)
base_r_squared_18 <- summary(base_model_18)$r.squared

results_list_18[[i]] <- list()

for (ii in seq_along(extensions_18)) {
    # Combine the base formula and the current extension
    extended_formula_18 <- update(base_formula_18, paste("~ . +", paste(extensions_18[[ii]], collapse = " + ")))
    
    # Fit the extended model
    extended_model_18 <- lm(extended_formula_18, data = data_18)

    # Get the R-squared for the extended model
    extended_r_squared_18 <- summary(extended_model_18)$r.squared

    # Calculate the incremental R-squared
    delta_r_squared_18 <- extended_r_squared_18 - base_r_squared_18
    
    # Likelihood Ratio Test
    LRT_18 <- anova(base_model_18, extended_model_18, test = "LRT")

    # Result List
    results_list_18[[i]][[ii]] <- list(
      base_r_squared = base_r_squared_18,
      extended_formula = extended_formula_18,
      extended_model_sumry = summary(extended_model_18),
      extended_r_squared = extended_r_squared_18,
      delta_r_squared = delta_r_squared_18,
      LRT = LRT_18)
  
    #Resetting base formula to last extension
    base_formula_18 <- extended_formula_18
    base_model_18 <- lm(base_formula_18, data = data_18)
    base_r_squared_18 <- summary(base_model_18)$r.squared
  }

}

results_list_18

```

# Descriptive Statistics (Used for both data sets)
```{r}

#2016 (and 2018 where applicable)

#Means and Standard Deviations (with weighted option), and medians
weighted_mean_sd(Survey_ICA_16$Q32_Avg, Survey_ICA_16$WEIGHT)

with(Survey_ICA_16[Survey_ICA_16$GENDER == 1, ], weighted_mean_sd(Q19, WEIGHT)) #Male Hesitancy
with(Survey_ICA_16[Survey_ICA_16$GENDER == 2, ], weighted_mean_sd(Q19, WEIGHT)) #Female Hesitancy

#Healthcare Recommendation Response
#2016
#Weighted counts
count(x = Survey_ICA_16, Q7, wt = WEIGHT)

#2018
table(Survey_ICA_18$Q7)/sum(table(Survey_ICA_18$Q7))

#Median (Education shown)
c("Median:", median(Survey_ICA_16$EDUC))


#2018 Mean and Standard Deviation (unweighted; Intention shown)
c("mean", mean(Survey_ICA_18$Q9, na.rm = TRUE))
c("sd", sd(Survey_ICA_18$Q9, na.rm = TRUE))

table(Survey_ICA_18$Q7)/sum(table(Survey_ICA_18$Q7))


###Cronbach's Alpha###

#Vaccine Confidence
cronbach.alpha(na.omit(Survey_ICA_16[,c("Q20_A","Q20_B","Q20_C")]))

#HDMP
cronbach.alpha(na.omit(Survey_ICA_16[,c( "Q16_A", "Q16_E", "Q16_B", "Q16_F", "Q16_D", "Q16_H")]))
#2018: "Q16_A", "Q16_F", "Q16_B", "Q16_G", "Q16_E", "Q16_I"
#2016: "Q16_A", "Q16_E", "Q16_B", "Q16_F", "Q16_D", "Q16_H"

#Perceived Flu Knowledge
cronbach.alpha(na.omit(Survey_ICA_18[,c("Q32_A","Q32_B","Q32_C", "Q32_D")]))
cronbach.alpha(na.omit(Survey_ICA_16[,c("Q32_A","Q32_B","Q32_C", "Q32_D")]))


#t-tests

#2016 - GENDER
design <- svydesign(ids = ~1, weights = ~WEIGHT, data = Survey_ICA_16)
svyttest(Q19 ~ factor(GENDER), design)
#2018 - GENDER
design <- svydesign(ids = ~1, data = Survey_ICA_18)
svyttest(Q19 ~ factor(GENDER), design)

```

# Scatter Plot Vaccine Confidence vs. Vaccine Hesitancy 
```{r}

#Sub-setting vaccine confidence and vaccine hesitancy
#2016
sub_2016 <- Survey_ICA_16 %>%
 dplyr::select(Q19, Q20_Avg) %>%
  mutate(Year = rep(2016, dim(Survey_ICA_16)[1]))
#2018
sub_2018 <- Survey_ICA_18 %>%
 dplyr::select(Q19, Q20_Avg) %>%
  mutate(Year = rep(2018, dim(Survey_ICA_18)[1]))

#Calculating Spearman correlation With p-values
cor.test(sub_2016$Q19, sub_2016$Q20_Avg, method = 'spearman') #2016
cor.test(sub_2018$Q19, sub_2018$Q20_Avg, method = 'spearman') #2018

#Plot
plot_image  <- ggplot(Combined_Data_16_18, aes(x = Q19, y = Q20_Avg))+
  geom_point(aes(shape = factor(Year), color = factor(Year)), alpha = 0.6) +
  geom_jitter(aes(shape = factor(Year), color = factor(Year)), width = 0.2, height = 0, alpha = 0.3) + 
  geom_smooth(aes(color = factor(Year)), method = "lm", se = FALSE) +
  scale_shape_manual(values = c("2016" = 16, "2018" = 17)) +  # 16: dot, 17: triangle
  scale_color_manual(values = c("2016" = "black", "2018" = "blue")) +
    labs(
    title = "Vaccine Hesitancy vs. Vaccine Confidence",
    subtitle = "With Linear Regression Lines and Spearman Correlations",
    x = "Vaccine Hesitancy Score ",
    y = "Average Vaccine Confidence Score",
    color = "Year", 
    shape = "Year"
  ) + theme_minimal()

plot_image

```






